---
title: "Picking a Language for Server-Side Programming (Working Draft)"
author: "Janos Brezniczky"
date: "January 2017"
output:
  md_document:
    toc: true
---

```{r echo=FALSE}
# setwd("/media/janca/Code/Prog/Github Analysis/analytics-and-hadoop-trends/github-trends/")
source("plotting.R")

load("intermediate_server/results.RData")
github.results = results
load("intermediate_server/SO_results.RData")
SO.results = results
rm(results)

color.table = get.color.table(c(github.results, SO.results))
```

# Picking a Language for Server-Side Programming (Working Draft)

_Janos Brezniczky_

_January, 2017_

## Introduction

I don't have much to do with server-side programming these days, so I'll let the charts speak for themselves, mostly.

The [Wikipedia Diffusion of innovations](https://en.wikipedia.org/wiki/Diffusion_of_innovations) article may give some ideas what curve types we see on the charts, however it is difficult to decide the impregnation of which area(s) are described by each curve from merely these graphs, and to find out which of these, if any, have already started to reach into the flattening out (late majority) phase.

## Methodology

The analysis is based on the [GitHub](http://github.com) public repository statistics.

To gather the data, a GitHub API client script was used to obtain weekly statistics, 
then a moving average was applied to get values smoothed over 53-week windows to get 
visually comprehensible charts and compensate for seasonality effects. Where the data does 
not allow for the entire filtering window width, the uncertainity (variance) increases, this 
is represented by fainter (more transparent) colors, as seen on the edges.

## Keyword #1: "Server"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$Server, SO.results$Server, "Server")
```

## Keyword #2: "Web Server"

```{r echo=FALSE, fig.width=10}
smooth.plot(github.results$`Web Server`, "Web Server")
```

## Keyword #3: "Service"

```{r echo=FALSE, fig.width=10}
smooth.plot(github.results$Service, "Service")
```

## Keyword #4: "Microservice"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$Microservice, SO.results$Microservice, "Microservice")
```

## Future Work

A problematic aspect is that repositories are normally multi-lingual, thus "intersection sets" will need to be checked on and accounted for.
An interesting thing can be the synchronous jump in PHP and JavaScript around early 2016.

A very fundamental issue is just by these charts it is impossible to compare the actual popularity of "servers" e.g. in Python vs. R, as the word "Analysis" just happens to appear in any text at times, and Python has a much larger user base, hence a higher volume of standard activity (not closely related to analytics).

## Conclusion

No conclusion here for now - choose whatever best fits.
