---
title: "Picking a Language for Analytics and Machine Learning (Working Draft)"
author: "Janos Brezniczky"
date: "January 2016"
output:
  md_document:
    toc: true
---

```{r echo=FALSE}
# setwd("/media/janca/Code/Prog/Github Analysis/analytics-and-hadoop-trends/github-trends/")
source("plotting.R")

load("intermediate/results.RData")
```

# Picking a Language for Analytics and Machine Learning (Working Draft)

_Janos Brezniczky_

_January, 2016_

## Executive Conjecture

Python.

## Introduction

Which language to choose for analytics? It can be hard to tell.

The competition of programming langauges is already a long story and continues to stretch on.
Judging the topic is difficult, subjective and is prone to statistical methodological issues.

This document is aimed at getting some indication of the trends, but by no means to come up with a definitive result or judgement over any debate.

The present version is in an exploratory/draft status, and attempts to achieve even less - to be a starting point for the above.

To gather the data, a GitHub API client script was used to obtain weekly statistics, then a moving average was applied to get values smoothed over 53-week windows to get visually comprehensible charts and compensate for seasonality effects.
The R vs. 'Analysis' numbers clearly feature an outlier, it is yet to be looked into what event this corresponds with.

## Keyword #1: "Analysis"

The most interesting here is probably that the influence of R seems to plateau out.

```{r echo=FALSE, fig.width=10}
smooth.plot(results$Analysis, "Analysis")
```

For some, the likely surprise of the (currently fractional) year 2016 may also be that Python (at least in this aspect) seems to be overtaking the lead in being the prime vehicle for analytics from R. Beyond that, both are strong candidates and Python is obviously involved in more analysis related libraries than R (part of the perceived growth is that it is probably catching up), which is just one of the reasons to scratch our heads whether to accept such finding as a fact. However, probably not many are that surprised, and this is just what it is. The R ecosystem needs to pull itself together quick if it wants to achieve more than to see the Python train passing. Python seems to be in the early adoption/early majority phase, which signifies a prospect for heavy future growth in this area. (See [Wikipedia](https://en.wikipedia.org/wiki/Diffusion_of_innovations).)

(Note that GitHub is a distorted representation - a large proportion of the repositories is simple coursework... but also mind that coursework is a good predictor of future preference.)

## Keyword #2: "Spark"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$Spark, "Spark")
```

Scala steadily seems to be the major player at Spark.
SparkR is in a pre-explosion state, if anywhere - and R didn't make it to the top 10 in any year.

However, looking at ItJobsWatch provides a slightly different result at the time writing, at least in the London context: Java and Apache Spark are more often colocated in job advertisements than Scala and Apache Spark. But! Java is big (30%) otherwise, some Spark users and thus teams will have a Hadoop background with a carried over Java preference, and not sure how much of its slight dominance would be left once we corrected for that.

See http://www.itjobswatch.co.uk/jobs/london/apache%20spark.do for the current status in case of interest.

Similarly GitHub for the past 6 months indicates a trendshift, too - however due to potential seasonality issues and overwhelming Christmas laziness that's not been presented here. Next version.


## Keyword #3: "Machine Learning"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Machine Learning`, "Machine Learning")
```

I think this one needs no explanation. In my experience, GitHub repos are quite often related to online learning, and the conclusion to draw is that Python is the de facto standard of choice for teaching machine learning.
One remarkable aspect is perhaps that the growth is beautfully exponential - it is typically an early phase of an adoption curve, again.

## Keyword #4: "Deep Learning"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Deep Learning`, "Deep Learning")
```


## Keyword #5: "Big Data"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Big Data`, "Big Data")
```

## Keyword #6: "AWS"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$AWS, "AWS")
```

## Keyword #7: "Data Science"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Data Science`, "Data Science")
```

## Keyword #8: "Science"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Science`, "Science")
```

## Keyword #9: "Data"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Data`, "Data")
```

## +1: "Kaggle"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$`Kaggle`, "Kaggle")
```

## +2: "Coursera"

```{r echo=FALSE, fig.width=10}
smooth.plot(results$Coursera, "Coursera")
```

GitHub, the most popular free source source code repository service, is the vehicle for many assignments on Coursera, which then is the most popular MOOC, and as such, heavily affects even a massive vehicle like GitHub.
From the above chart it already gets clear that these courses are at the heart of R's popularity on GitHub, and are accountable for at least one significant growth period (starting around mid-2013 +- 6 months) in the use of R. This probably was the [Data Science Specialization](https://www.coursera.org/specializations/jhu-data-science).

However, this impact is not necessarily a bad thing from the perspective of the analysis - future analysts are trained in the present, and so does their current preference drive future use.

## Future Work

JavaScript and HTML have notoriously been omitted from these statistics.
These are considered the presentation or output in this analysis and thus not the 'active ingredient'. While JavaScript is included on the AWS charts, it is difficult to tell to what extent it was useful to do so.
However, the means of presentation is an important factor of an analysis, and its merits will be re-evaluated somehow.
Another, strongly related and problematic aspect is that repositories are normally multi-lingual, thus "intersection sets" will need to be checked on and accounted for. A similar and more worrying problem is that multi-word search phrases, like ["Big Data"] are not actually treated as a single literal, but probably as ["Big" + "Data"] (i.e. or'ed), thus giving distorted results.

## Conclusion

Python.

At the minute, it seems, if you only choose one programming language for analytics (however over-idealistic that situation may be), or about to choose your first one, it is Python for the win.
It appears to be of significant presence (transferrable) or sheer dominance (hard to dismiss) in every single aspect examined above.
And you won't be alone with using it.
There are lots of things to work out in such an exploding ecosystem, and in the best case you're part of the happenings.
