---
title: "Picking a Language for Server-Side Programming (Working Draft)"
author: "Janos Brezniczky"
date: "January 2017"
output:
  md_document:
    toc: true
---

```{r echo=FALSE}
# setwd("/media/janca/Code/Prog/Github Analysis/analytics-and-hadoop-trends/github-trends/")
source("plotting.R")

load("intermediate_server/results.RData")
github.results = results
load("intermediate_server/SO_results.RData")
SO.results = results
rm(results)

color.table = get.color.table(c(github.results, SO.results))
series.order = get.series.order(c(github.results, SO.results))
```

# Picking a Language for Server-Side Programming (Working Draft)

_Janos Brezniczky_

_March, 2017_

## Introduction

I don't have much to do with server-side programming these days, so I'll
let the charts speak for themselves, mostly.

The [Wikipedia Diffusion of
innovations](https://en.wikipedia.org/wiki/Diffusion_of_innovations)
article may give some ideas what curve types (or the superposition of those)
we can see on the charts, however it is difficult to decide the impregnation 
of which area(s) are described by each curve from merely these graphs, 
and to find out which of these, if any, have already started to reach 
into the flattening out (late majority) phase.

## Methodology

The document is based on the [GitHub](http://github.com) public
repository and [StackOverflow](http://stackoverflow.com) question 
statistics.

The data is weekly statistics, obtained from both sites via their public API.

A moving average filter is applied next, to get values smoothed 
over 53-weekwindows to get visually comprehensible charts and 
compensate for seasonality effects. Where the data does not 
allow for the entire filtering window width, the uncertainity (variance)
increases, this is represented by fainter (more transparent) colors, 
as seen near the edges.

## Keyword #1: "Server"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$Server, SO.results$Server, "Server")
```

## Keyword #2: "Web Server"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$`Web Server`, SO.results$`Web Server`, "Web Server")
```

## Keyword #3: "Service"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$Service, SO.results$Service, "Service")
```

## Keyword #4: "Microservice"

```{r echo=FALSE, fig.width=10}
double.plot(github.results$Microservice, SO.results$Microservice, "Microservice")
```

## Future Work

A problematic aspect is that repositories are normally multi-lingual, thus "intersection sets" will need to be checked on and accounted for.
An interesting thing can be the synchronous jump in PHP and JavaScript around early 2016.

A very fundamental issue is just by these charts it is not possible to accurately compare the actual popularity of "servers" e.g. in Python vs. R, as the word "server" just happens to appear in any text at times, and Python has a much larger user base, hence a higher volume of standard activity (not closely related to analytics).

## Conclusion

No conclusion here for now - choose whatever best fits.
