---
title: "Picking a Language for Server-Side Programming (Working Draft)"
author: "Janos Brezniczky"
date: "April 2017"
output:
  md_document:
    toc: true
---

```{r echo=FALSE}
# setwd("/media/janca/Code/Prog/Github Analysis/analytics-and-hadoop-trends/github-trends/")
source("plotting.R")

load("intermediate_server/results.RData")
github.results = results
load("intermediate_server/SO_results.RData")
SO.results = results
rm(results)

color.table = get.color.table(c(github.results, SO.results))
series.order = get.series.order(c(github.results, SO.results))
```

# Picking a Language for Server-Side Programming (Working Draft)

_Janos Brezniczky_

_April, 2017_

## Introduction

I don't have much to do with server-side programming these days, so I'll
let the charts speak for themselves, mostly.

The [Wikipedia Diffusion of
innovations](https://en.wikipedia.org/wiki/Diffusion_of_innovations)
article may give some ideas what curve types (or the superposition of those)
we can see on the charts, however it is difficult to decide the impregnation 
of which area(s) are described by each curve from merely these graphs, 
and to find out which of these, if any, have already started to reach 
into the flattening out (late majority) phase.

## Methodology

The document is based on the [GitHub](http://github.com) public
repository and [StackOverflow](http://stackoverflow.com) question 
statistics.

The data is weekly statistics, obtained from both sites via their public API.

A moving average filter is applied next, to get values smoothed 
over 53-week windows to get visually comprehensible charts and 
compensate for seasonality effects. Where the data does not 
allow for the entire filtering window width, the uncertainity (variance)
increases, this is represented by fainter (more transparent) colors, 
as seen near the edges.

## Keyword #1: "Server"

```{r "k1", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$Server, SO.results$Server, "Server")
```

Just guesses: What we can see is the interest in server programming seems to stagnate/decrease on StackOverflow. My guess here is that on the one hand, various stacks and the adoption of methodologies more commonplace elsewhere (such as OOP in JavaScript) probably have boosted productivity. On the other hand, servers themselves are less likely to be implemented ground up. I'd expect it is also less likely that the server itself is the product, or is written ground up, rather the application-specific features. It would be also nice to verify if, for instance, a few popular servers are becoming the standard. A quick check: as of 14/04/2017, about half of the SO Python Server questions are related to Django. A better documentation, questions already answered can also be related.

## Keyword #2: "Cloud"
```{r "k2", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$Cloud, SO.results$Cloud, "Cloud")
```

The interest in cloud-related topics seems to be on the incline.

While the increase in the latter doesn't fully explain the halt in the server topic growth, cloud can easily overtake the market from classic server hosting. Classic server languages definitely stopped gaining more interest - perhaps down to a better documentation, or as elsewhere mentioned (TODO: add link to prog.hu!), in the favour of newer-style, JS and Python based approaches.

## Keyword #3: "Serverless"
```{r "k3", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$Serverless, SO.results$Serverless, "Serverless")
```

Serverless is just my favourite candidate for the future. Spinning up a cluster takes time. Possibly the biggest drawback with cloud computing is having to wait for the response. If that is reduced, that's already great - but so is the lower need for operations-related duties.
This may enable people to focus on a smaller set of activities, and perform those to a higher quality. Whether an abstraction above e.g. AWS Lambda's is to come, is a further question. These requests are not meant to run infinitely long - but they are said to scale well. Operations in the cloud anyway are typically somewhat fragmented so that finishing an interrupted operation becomes more effortless.

## Keyword #4: "Service"

```{r "k4", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$Service, SO.results$Service, "Service")
```

## Keyword #5: "Microservice"

```{r "k5", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$Microservice, SO.results$Microservice, "Microservice")
```

## Keyword #6: "Web Server"

```{r "k6", echo=FALSE, fig.width=10, cache=TRUE}
double.plot(github.results$`Web Server`, SO.results$`Web Server`, "Web Server")
```

## Future Work

A problematic aspect is that repositories are normally multi-lingual, thus "intersection sets" will need to be checked on and accounted for.
An interesting thing can be the synchronous jump in PHP and JavaScript around early 2016.

A very fundamental issue is just by these charts it is not possible to accurately compare the actual popularity of "servers" e.g. in Python vs. R, as the word "server" just happens to appear in any text at times, and Python has a much larger user base, hence a higher volume of standard activity (not closely related to analytics).

## Conclusion

No conclusion here for now - choose whatever best fits.
