---
title: "Picking a Language for Server-Side Programming (Working Draft)"
author: "Janca"
date: "23 December 2016"
output:
  md_document:
    toc: true
---

```{r echo=FALSE}

# setwd("/media/janca/Code/Prog/Github Analysis/analytics-and-hadoop-trends")

# read in data
library(knitr)

all.cols = c("blue", "darkgreen", "red", "green", "black", "yellow", "purple", "brown", "orange")

# do not forget the k-filter's radius
start.date = as.Date("2008-01-01") + 22 * 7

load("intermediate_server/results.RData")

smooth.plot = function(raw.values, main) {
  # 53 wide kernel filter is used, close enough to
  # a yearly basis, which is relatively robust against
  # seasonality based fluctuations
  k1 = kernel("daniell", 26)

  # keep recurring items at well-defined locations to
  # enhance readability and colour coding based consistency
  move.to = function(n, item, target.idx) {
    idx = which(n == item)
    if (length(idx) > 0) {
      n[idx] = n[target.idx]
      n[target.idx] = item
    }
    return(n)
  }
  
  n = names(raw.values)
  n = move.to(n, "Java", 1)
  n = move.to(n, "C#", 2)
  n = move.to(n, "go", 3)
  n = move.to(n, "Python", 4)
  raw.values = raw.values[n]
  
  values = list()
  for(name in names(raw.values)) {
    values[[name]] = 52 * kernapply(raw.values[[name]], k1)
  }
  
  ymax = max(unlist(sapply(values, max)))
  ymin = min(unlist(sapply(values, min)))
  
  xs = start.date + 0:(length(values[[1]]) - 1) * 7
  mx.values = do.call(cbind, values)
  cols = all.cols[1:ncol(mx.values)]
  matplot(xs,
          mx.values, ylim=c(ymin, ymax), type="l",
          ylab="", xlab="", xaxt="n",
          main=sprintf("'%s' repositories created on GitHub per year (est.)", main),
          # legend=rownames(values),
          # beside=TRUE,
          # args.legend=(x=ncol(counts) + 3),
          col=cols, lty=1, las=2, bty="n")
  # xs = xs[1 + round((0 : 7) * (length(xs) - 1) / 7)]
  xs = c(as.Date(c("2008-06-15", "2009-06-15", "2010-06-15", "2011-06-15", "2012-06-15", "2013-06-15", "2014-06-15", "2015-06-15", "2016-06-15")))
  timelabels=format(xs,"%Y-%m")  #%H:%M
  axis(1, at=xs, labels=timelabels, las=2, xlim=c(min(xs), max(xs)))
  legend(xs[1] - 100, max(do.call(c, values)), legend=names(values), col=cols, bty="n",
         lwd=2, lty=1)
}
```

# Picking a Language for Server-Side Programming (Working Draft)

## Introduction

I don't have much to do with server-side programming these days, so I'll let the charts speak for themselves, mostly.

The method was to search for keywords using a GitHub client script over weekly intervals, starting from 2008-01-01, and then using a 53 week moving average to plot some comprehensibly smooth charts and somewhat compensate for likely seasonality effects.

The [Wikipedia Diffusion of innovations](https://en.wikipedia.org/wiki/Diffusion_of_innovations) article may give some ideas what curve types we see on the charts, however it is hard to decide the impregnation of what area(s) are described by each curve from merely these graphs, and to find out which of these, if any, have already started to reach into the flattening out (late majority) phase.

## Keyword #1: "Server"

```{r echo=FALSE}
smooth.plot(results$Server, "Server")
```

## Keyword #1: "Web Server"

```{r echo=FALSE}
smooth.plot(results$`Web Server`, "Web Server")
```

## Keyword #2: "Service"

```{r echo=FALSE}
smooth.plot(results$Service, "Service")
```

## Keyword #3: "Microservice"

```{r echo=FALSE}
smooth.plot(results$`Microservice`, "Microservice")
```

## Future work

A problematic aspect is that repositories are normally multi-lingual, thus "intersection sets" will need to be checked on and accounted for.

## Conclusion

No conclusion here - choose whatever best fits.

Note: PHP and JavaScript (or derivatives, like Node.JS) have probably been unreasonably left out from the above comparisons and will be added soon where relevant.

